Writing a reproducible paper with RStudio and Quarto

### 2023-03-05 paper.qmd
edits.

---
title: "Writing a reproducible paper with RStudio and Quarto"
author:
  - name: Paul C. Bauer
    orcid: 0000-0002-8382-9724
    email: mail@paulcbauer.de
    affiliations:
      - name: Mannheim Centre for European Social Research
  - name: Camille Landesvatter
    orcid: 0000-0003-0156-5364
    email: camille.landesvatter@uni-mannheim.de
    affiliations:
      - name: Mannheim Centre for European Social Research
format:
  html: default
  pdf: default
date: 'last-modified'
date-format: "[First version:February  16, 2023<br>This version:] MMMM D, YYYY"
abstract: 'This paper provides a template to write a reproducible scientific paper with RStudio and Quarto. It falls into a series on other templates, namely on RMarkdown and pagedown that can be downloaded [here](https://osf.io/k8jhx/) [@Bauer2018-dl] and [here](https://osf.io/q395s) [@Bauer2021-dl]. Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s). We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto. We intend to update this file when we discover more convenient code. You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/).'
bibliography: references.bib
csl: american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute: 
  echo: false
  warning: false
  message: false
---

@@

* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.

@@

* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].

```{{r}}
install.packages(c("knitr", "kable", "kableExtra",
                   "stargazer", "modelsummary", "gt"))
```

@@

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
#| label: tbl-6
#| tbl-cap: "Linear regression"
library(stargazer)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
stargazer(M1, M2, M3,
          title = "Regression table with stargazer",
          label="tab2", 
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

@@

## R base graphs
Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.qmd` file. It's relatively simple to insert R base graphs as you can see in @fig-1.

@@

Below in @fig-2 we also show a ggplot2 example.

@@

```{r}
#| label: fig-2
#| fig-cap: "Miles per gallon according to the weight"
library(ggplot2)
plot <- ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) +
  geom_point() +
  labs(x="Agriculture", y = "Fertility", shape="Education") +
  theme_classic()
plot
```

>>>>

---
title: "Writing a reproducible paper with RStudio and Quarto"
author:
  - name: Paul C. Bauer
    orcid: 0000-0002-8382-9724
    email: mail@paulcbauer.de
    affiliations:
      - name: Mannheim Centre for European Social Research
  - name: Camille Landesvatter
    orcid: 0000-0003-0156-5364
    email: camille.landesvatter@uni-mannheim.de
    affiliations:
      - name: Mannheim Centre for European Social Research
format:
  html: default
  pdf:
    number-sections: true
    colorlinks: true
    keeptex: true
date: 'last-modified'
date-format: '[First version: February  16, 2023 <br>This version:] MMMM D, YYYY'
abstract: 'This paper provides a template to write a reproducible scientific paper with RStudio and Quarto. It falls into a series on other templates, namely on RMarkdown and pagedown that can be downloaded [here](https://osf.io/k8jhx/) [@Bauer2018-dl] and [here](https://osf.io/q395s) [@Bauer2021-dl]. Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s). We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto. We intend to update this file when we discover more convenient code. You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/).'
bibliography: references.bib
csl: american-sociological-association.csl
tbl-cap-location: top
number-sections: true
execute: 
  echo: false
  warning: false
  message: false
---

@@

* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. Quarto allows you to **automatize** such steps and/or avoid them.

@@

* ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @hlavac2013stargazer; @modelsummary; @gt].

```{{r}}
install.packages(c("knitr", "kableExtra",
                   "stargazer", "modelsummary", "gt"))
```

@@

```{r echo=TRUE, paged.print=FALSE, results="asis"}
#| label: tbl-6
#| tbl-cap: "Linear regression"
library(stargazer)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
stargazer(M1, M2, M3,
          title = "Regression table with stargazer",
          label="tab2", 
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

@@

## R base graphs
A fully reproducible manuscript would create diagrams directly in the `.qmd` file and insert them in the appropriate place. It's relatively simple to insert R base graphs as you can see in @fig-1.

@@

Below in @fig-2 we also show a example witrh R's `ggplot2` package.

@@

```{r fig-2, fig.align="center", fig.pos="H", fig.width=6, fig.height=3}
#| label: fig-2
#| fig-cap: "Fertility and Agriculture by Education"
library(ggplot2)
plot <- ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) +
  geom_point() +
  labs(x="Agriculture", y = "Fertility", shape="Education") +
  theme_classic()
plot
```
